Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-05-21 PAPER.tex
edited jeffreys prior appendix according to Adrian edits

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis (one bin per injection level), and, assuming we are accurately sampling a Gaussian distribution, we smooth the distribution of $\widehat{P}_{\rm out}$ values by fitting Gaussians for each bin based on its mean and variance (and normalize them). Stitching all of them together results in a 2-dimensional transfer function --- the likelihood function in Bayes' rule, namely $\mathcal{L} (  \widehat{P}_{\rm out} | P_{\rm in})$. We then have a choice for our prior, $p(P_{\rm in})$, and we choose to invoke a Jeffreys prior (\citealt{jaynes1968}) because it is a true uninformative prior. For a derivation and more details about the Jeffreys prior used in our analysis, see Appendix \ref{sec:jeffreys}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.pdf}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $\widehat{P}_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Power spectra values (black points) are generated for $20$ realizations of $\textbf{e}$ per signal injection level. Assuming our $\widehat{P}_{\rm out}$ values sample a Gaussian distribution, we fit Gaussians to each injection level based on the mean and variance of the simulation outputs. This entire likelihood function is then multiplied by a Jeffreys prior for $p(P_{\rm in}$), with the final result shown as the colored 
heat-maps on top of the points. Two cases are displayed: empirically estimated inverse covariance weighted PAPER-64 data (left) and uniform-weighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{P}}_{x}$, from which a posterior distribution for the signal is extracted. From these plots, it is clear that the weighted case results in $\sim4$ orders of magnitude of signal loss at the data-only power spectrum value, whereas the uniform-weighted case does 
not exhibit loss. The general shape of these transfer functions are also shown by the black curves in Figure \ref{fig:sigloss_terms} for comparison.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

\begin{equation}
\label{eq:logprob}
\mathcal{L} = log \, p(\widehat{P}_{\rm out} | P_{\rm in}),
\end{equation}

@@

For a single injection amplitude, our bootstrapped $\widehat{P}_{\rm out}$ values sample a Gaussian distribution. Simplifying our notation so that $x = P_{\rm in}$ and $y = \widehat{P}_{\rm out}$:

@@

\noindent where $\sigma$ is the standard deviation of $\widehat{P}_{\rm out}$ and $\bar y$ is the mean of $\widehat{P}_{\rm out}$, and they are both functions of $P_{\rm in}$. Plugging in Equation \eqref{eq:prob} into Equation \eqref{eq:logprob} becomes:

\begin{equation}
\label{eq:logprob_full}
\mathcal{L} = -\, log \, \sqrt{2\pi} - log\, \sigma - \frac{1}{2}\Big(\frac{y-\bar y}{\sigma}\Big)^{2}. 
\end{equation}

Next, the quantity inside the expectation value of Equation \eqref{eq:jeffreys} can be computed by taking the derivative of Equation \eqref{eq:logprob_full} with respect to $P_{\rm in}$, or $x$:

@@

When we simulate our full injection framework as in Section \ref{sec:Practice}, we can easily fit analytic functions to the mean and standard deviations of $\widehat{P}_{\rm out}$ ($\bar y$ and $\sigma$) as functions of $P_{\rm in}$. An example of the typical shape of these functions for the PAPER-64 analysis is shown in Figure \ref{fig:jeffreys1}, though we note in practice we fit solutions for every $k$-value and simulation independently.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/jeffrey_fits.pdf}
	\caption{An illustrative example (for the PAPER-64 analysis using uniform weighting and $k=0.393$\,$h$ Mpc$^{-1}$) of how the mean of $P_{\rm out}$ (left) and standard deviation of $P_{\rm out}$ (right) behave as a function of $P_{\rm in}$. Polynomials are fit to each (red) to describe how $\bar y$ and $\sigma$ evolve with $x$ (injection level), respectively, for the computation of the Jeffreys prior as defined in Equation \eqref{eq:jeffreys_final}.}
	\label{fig:jeffreys1}
\end{figure}

We also show the typical shape of the Jeffreys prior used in our analysis in Figure \ref{fig:jeffreys2}, as computed by Equation \eqref{eq:jeffreys_final}. Most noticeably, it is not constant with $P_{\rm in}$, meaning a uniform prior, which is often assumed to be an uninformative prior, is actually informative. Therefore, due to its objective nature we choose to use a Jeffreys prior in our analysis, multiplying our likelihood functions by Equation \eqref{eq:jeffreys_final} before computing posterior distributions.

\begin{figure}
	\centering
	\includegraphics[width=10cm]{plots/jeffrey_prior.pdf}
	\caption{An example of the typical Jeffreys prior shape for the PAPER-64 analysis. Most noticeably, it is not constant with $P_{\rm in}$, meaning a uniform prior would be an informative prior.}
	\label{fig:jeffreys2}
\end{figure}

>>>>

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis (one bin per injection level) and, since they are well-approximated by a Gaussian distribution in our numerical results, we smooth the distribution of $\widehat{P}_{\rm out}$ values by fitting Gaussians for each bin based on its mean and variance (and normalize them). Stitching all of them together results in a 2-dimensional transfer function --- the likelihood function in Bayes' rule, namely $\mathcal{L} (  \widehat{P}_{\rm out} | P_{\rm in})$. We then have a choice for our prior, $p(P_{\rm in})$, and we choose to invoke a Jeffreys prior (\citealt{jaynes1968}) because it is a true uninformative prior. For a derivation and more details about the Jeffreys prior used in our analysis, see Appendix \ref{sec:jeffreys}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.pdf}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $\widehat{P}_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Power spectra values (black points) are generated for $20$ realizations of $\textbf{e}$ per signal injection level. Since our $\widehat{P}_{\rm out}$ values are well-approximated by a Gaussian distribution, we fit Gaussians to each injection level based on the mean and variance of the simulation outputs. This entire likelihood function is then multiplied by a Jeffreys prior for $p(P_{\rm in}$), with the final result shown as the colored 
heat-maps on top of the points. Two cases are displayed: empirically estimated inverse covariance weighted PAPER-64 data (left) and uniform-weighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{P}}_{x}$, from which a posterior distribution for the signal is extracted. From these plots, it is clear that the weighted case results in $\sim4$ orders of magnitude of signal loss at the data-only power spectrum value, whereas the uniform-weighted case does 
not exhibit loss. The general shape of these transfer functions are also shown by the black curves in Figure \ref{fig:sigloss_terms} for comparison.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

\begin{equation}
\label{eq:logprob}
\mathcal{L} = \mathrm{ln} \, p(\widehat{P}_{\rm out} | P_{\rm in}),
\end{equation}

@@

For a single injection amplitude, our bootstrapped $\widehat{P}_{\rm out}$ values are well-approximated by a Gaussian distribution. Simplifying our notation so that $x = P_{\rm in}$ and $y = \widehat{P}_{\rm out}$:

@@

\noindent where $\sigma$ is the standard deviation of $\widehat{P}_{\rm out}$ and $\bar y$ is the mean of $\widehat{P}_{\rm out}$, and they are both functions of $P_{\rm in}$. Using Equations \eqref{eq:prob} and \eqref{eq:logprob}, the quantity inside the expectation value of Equation \eqref{eq:jeffreys} becomes:

%\begin{equation}
%\label{eq:logprob_full}
%\mathcal{L} = -\, log \, \sqrt{2\pi} - log\, \sigma - \frac{1}{2}\Big(\frac{y-\bar y}{\sigma}\Big)^{2}. 
%\end{equation}

%Next, the quantity inside the expectation value of Equation \eqref{eq:jeffreys} can be computed by taking the derivative of %Equation \eqref{eq:logprob_full} with respect to $P_{\rm in}$, or $x$:

@@

When we simulate our full injection framework as in Section \ref{sec:Practice}, we sample $P_{\rm in}$ values that range from $\sim$$10^{5}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$ to $\sim$10$^{11}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, so we note that the prior is set to zero outside those regions. For the injections that we do sample, we can easily fit analytic functions to the mean and standard deviations of $\widehat{P}_{\rm out}$ ($\bar y$ and $\sigma$) as functions of $P_{\rm in}$. An example of the typical shape of these functions for the PAPER-64 analysis is shown in Figure \ref{fig:jeffreys1}, though we note in practice we fit solutions for every $k$-value and simulation independently.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/jeffrey_fits.pdf}
	\caption{An illustrative example (for the PAPER-64 analysis using uniform weighting and $k=0.393$\,$h$ Mpc$^{-1}$) of how the mean of $P_{\rm out}$ (left) and standard deviation of $P_{\rm out}$ (right) behave as a function of $P_{\rm in}$. Polynomials are fit to each (red) to describe how $\bar y$ and $\sigma$ evolve with $x$ (injection level), respectively, for the computation of the Jeffreys prior as defined in Equation \eqref{eq:jeffreys_final}. The polynomial fits for this example are $y = (-5.1 \times 10^{-15})x^{2} + x + (1.5 \times 10^{7})$ and $y = (5.0 \times 10^{-13})x^{2} + 0.2 x + 10^{7}$ for $\bar y$ and $\sigma$, respectively.}
	\label{fig:jeffreys1}
\end{figure}

We also show the typical shape of the Jeffreys prior used in our analysis in Figure \ref{fig:jeffreys2}, as computed by Equation \eqref{eq:jeffreys_final}. Most noticeably, it is not constant with $P_{\rm in}$, meaning a uniform prior, which is often used for simplicity, is informative in our application. Therefore, due to its objective nature we choose to use a Jeffreys prior in our analysis, multiplying our likelihood functions by Equation \eqref{eq:jeffreys_final} before computing posterior distributions.

\begin{figure}
	\centering
	\includegraphics[width=10cm]{plots/jeffrey_prior.pdf}
	\caption{An example of the typical Jeffreys prior shape for the PAPER-64 analysis as computed by Equation \eqref{eq:jeffreys_final}. Most noticeably, it is not constant with $P_{\rm in}$, meaning a uniform prior would be an informative prior.}
	\label{fig:jeffreys2}
\end{figure}
